---
title: "Lab2"
author: "Ahsanul Choudhury"
date: "September 11, 2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, include=TRUE}
library(IS606)
library(ggplot2)
```

```{r load-data, eval=TRUE}
load("more/kobe.RData")
head(kobe)
```

##Using calc_streak, compute the streak lengths of sim_basket.

##1.Describe the distribution of streak lengths. What is the typical streak length for this simulated independent shooter with a 45% shooting percentage? How long is the player's longest streak of baskets in 133 shots?

##2.If you were to run the simulation of the independent shooter a second time, how would you expect its streak distribution to compare to the distribution from the question above? Exactly the same? Somewhat similar? Totally different? Explain your reasoning.

##3.How does Kobe Bryant's distribution of streak lengths compare to the distribution of streak lengths for the simulated shooter? Using this comparison, do you have evidence that the hot hand model fits Kobe's shooting patterns? Explain.

```{r sim-basket1, eval=TRUE}
outcomes <- c("H", "M")
sim_basket1 <- sample(outcomes, size = 133, replace = TRUE, prob=c(0.45, 0.55))
```

```{r kobe_streak, eval=TRUE}
kobe_streak <- calc_streak(kobe$basket)
kobe_streak
table(kobe_streak)
summary(kobe_streak)
```

```{r sim_streak1, eval=TRUE}
sim_streak1 <- calc_streak(sim_basket1)
sim_streak1
table(sim_streak1)
summary(sim_streak1)
```

```{r multiplot_function, eval=TRUE}
multiplot <- function(..., plotlist = NULL, file, cols = 1, layout = NULL) {
  require(grid)

  plots <- c(list(...), plotlist)

  numPlots = length(plots)

  if (is.null(layout)) {
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                    ncol = cols, nrow = ceiling(numPlots/cols))
  }

  if (numPlots == 1) {
    print(plots[[1]])

  } else {
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))

    for (i in 1:numPlots) {
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))

      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}
```
##Reference:Multiple graphs on one page (ggplot2). (n.d.). Retrieved September 11, 2016, from http://www.cookbook-r.com/Graphs/Multiple_graphs_on_one_page_(ggplot2)/


```{r plot_sim1, eval=TRUE}
p2 <- qplot(sim_streak1, geom="histogram", binwidth = 1,
      main = "Sim1",
      xlab = "Streak",ylab = "Frequency",
      fill=I("lightgreen"),col=I("red"))

p2
```

##Answer 1. The distribution of the independent shooter is skwed to the right with typical streak length of 0. Longest streak is:

```{r sim1_lingest, eval=TRUE}
max(sim_streak1)
```

##Answer 2.
```{r sim-basket2, eval=TRUE}
outcomes <- c("H", "M")
sim_basket2 <- sample(outcomes, size = 133, replace = TRUE, prob=c(0.45, 0.55))
```


```{r sim_streak2, eval=TRUE}
sim_streak2 <- calc_streak(sim_basket2)
sim_streak1
table(sim_streak2)
summary(sim_streak2)
```

```{r plot_sim1vssim2, eval=TRUE}
p3 <- qplot(sim_streak2, geom="histogram", binwidth = 1,
      main = "Sim2",
      xlab = "Streak",ylab = "Frequency",
      fill=I("yellow"),col=I("red"))
multiplot(p2, p3, cols = 2)
```

##If we run the simulation again we can see from the plots above they are some what similar, skewness stays the same, typical streak length stays at 0 but the maximum length changes.

##Answer 3.
```{r plot_kobe, eval=TRUE}
p1 <- qplot(kobe_streak, geom="histogram", binwidth = 1,
      main = "Kobe",
      xlab = "Streak",ylab = "Frequency",
      fill=I("lightblue"),col=I("red"))
multiplot(p1, p2, p3, cols = 2)
```

##Kobe Bryant's distribution of streak lengths is similar to the simulated independent shooter, skwed to the right and typical streak length of 0. We do not have enough evidence to approve or deny that the hot hand model fits Kobe's shooting patterns but the results indicates that Koby's shooting pattern is an independent event rather than a denendent hot hand event.
